Triangulating iconicity: Coding analysis

### 2024-04-03 legacy/TI_03_ratings.Rmd
archive earlier, more fragmentary .Rmd's in /legacy/ for clarity

```{r}
# load corrected data
d <- read_xlsx("data\\ideophones_rated.xlsx")
d.means <- read_xlsx("data\\ideophones_rated_means.xlsx")

# get means and standard deviations across all ideophones and for each language and category
mean(d$rating)
  # 2.948339
sd(d$rating)
  # 1.297106
sumcat <- ddply(d,~category,summarise,mean=mean(rating),sd=sd(rating))
sumlang <- ddply(d,~language,summarise,mean=mean(rating),sd=sd(rating))

@@

```{r}
#create copy of data including the category "Other"
d.other <- read_xlsx("data\\ideophones_rated.xlsx")
d.m.other <- read_xlsx("data\\ideophones_rated_means.xlsx")

#remove category "Other" from d and d.means
d <- d[!grepl("Other",d$category),]
d.means <- d.means[!grepl("Other",d.means$category),]

# make new summary for language (without the Japanese "other" ideophones)
sumlang2 <- ddply(d,~language,summarise,mean=mean(rating),sd=sd(rating))
```

@@

```{r}
#convert some columns to factors
cols <- c("category","language","study")
d[cols] <- lapply(d[cols], factor)
d.means[cols] <- lapply(d.means[cols], factor)

#combined boxplot and scatterplot by category
ggplot(d.means, aes(x=reorder(category,-rating),y=rating,color=category)) +
  theme_tufte(base_size=16) +
  geom_half_boxplot(aes(middle=mean(rating)),show.legend=F) +
  geom_half_point(show.legend=F) +
  scale_colour_viridis_d(option="D",alpha=0.8) +
  xlab("category") + scale_x_discrete(labels = c("Sound","Motion","Shape", 
  "Texture","Colour/Visual"))
#ggsave("figures\\rating_category.png",height=5,width=7.5)

#same, but by language
ggplot(d.means, aes(x=reorder(language,-rating),y=rating,color=language)) +
  theme_tufte(base_size=16) +
  geom_half_boxplot(aes(middle=mean(rating)),show.legend=F) +
  geom_half_point(show.legend=F) +
  scale_colour_viridis_d(option="D",alpha=0.8) +
  xlab("category")
#ggsave("figures\\rating_language.png",height=5,width=7.5)

@@

```{r}
# check normality by visual inspection
ggqqplot(d, "rating", facet.by = "category")
  # all points fall approximately along the reference line, so normality can be assumed

@@

```{r}
summary(lm.cat <- lm(rating ~ category, d))

# pairwise comparisons with Tukey adjustments in the emmeans package
emm.cat <- emmeans(lm.cat, "category")
pairs(emm.cat)
pwpm(emm.cat)
```

@@

```{r}
summary(lm.lang <- lm(rating ~ language, d))

>>>>

```{r}
# load corrected data
d.allratings <- read_xlsx("data\\ideophones_rated.xlsx")
d.allratings.means <- read_xlsx("data\\ideophones_rated_means.xlsx")

# get means and standard deviations across all ideophones and for each language and category
mean(d.allratings$rating)
  # 2.948339
sd(d.allratings$rating)
  # 1.297106
sumcat <- ddply(d.allratings,~category,summarise,mean=mean(rating),sd=sd(rating))
sumlang <- ddply(d.allratings,~language,summarise,mean=mean(rating),sd=sd(rating))

@@

```{r}
#create copy of data including the category "Other"
d.allratings.other <- read_xlsx("data\\ideophones_rated.xlsx")
d.allratings.means.other <- read_xlsx("data\\ideophones_rated_means.xlsx")

#remove category "Other" from d and d.means
d.allratings <- d.allratings[!grepl("Other",d.allratings$category),]
d.allratings.means <- d.allratings.means[!grepl("Other",d.allratings.means$category),]

# make new summary for language (without the Japanese "other" ideophones)
sumlang2 <- ddply(d.allratings,~language,summarise,mean=mean(rating),sd=sd(rating))
```

@@

```{r}
#convert some columns to factors
cols <- c("category","language","study")
d.allratings[cols] <- lapply(d.allratings[cols], factor)
d.allratings.means[cols] <- lapply(d.allratings.means[cols], factor)

#combined boxplot and scatterplot by category
ggplot(d.allratings.means, aes(x=reorder(category,-rating),y=rating,color=category)) +
  theme_tufte(base_size=16) +
  geom_half_boxplot(aes(middle=mean(rating)),show.legend=F) +
  geom_half_point(show.legend=F) +
  scale_colour_viridis_d(option="D",alpha=0.8) +
  xlab("category") + scale_x_discrete(labels = c("Sound","Motion","Shape", 
  "Texture","Colour/Visual"))
#ggsave("figures\\rating_category.png",height=5,width=7.5)

#same, but by language
ggplot(d.allratings.means, aes(x=reorder(language,-rating),y=rating,color=language)) +
  theme_tufte(base_size=16) +
  geom_half_boxplot(aes(middle=mean(rating)),show.legend=F) +
  geom_half_point(show.legend=F) +
  scale_colour_viridis_d(option="D",alpha=0.8) +
  xlab("category")
#ggsave("figures\\rating_language.png",height=5,width=7.5)

@@

```{r}
# check normality by visual inspection
ggpubr::ggqqplot(d.allratings, "rating", facet.by = "category")
  # all points fall approximately along the reference line, so normality can be assumed

@@

```{r}
summary(lm.cat <- lm(rating ~ category, d.allratings))

# pairwise comparisons with Tukey adjustments in the emmeans package
emm.cat <- emmeans::emmeans(lm.cat, "category")
pairs(emm.cat)
pwpm(emm.cat)
```

@@

```{r}
summary(lm.lang <- lm(rating ~ language, d.allratings))

### 2024-04-03 legacy/TI_05_plotting_explorations.Rmd
archive earlier, more fragmentary .Rmd's in /legacy/ for clarity

# get ratings data
d.ratings <- read_xlsx("data\\ideophones_rated_means.xlsx") %>%
  dplyr::select(-category,-list,-item)
d <- left_join(d,d.ratings,by=c("filename","study","language"))

@@

# write this dataset as the fullest dataset
write.csv(d,file="data/ideophones_coded_guessed_rated.csv",fileEncoding = "UTF-8")
write_xlsx(d,path="data/ideophones_coded_guessed_rated.xlsx")

# also write a simplified copy for easy inspection in Excel (helpful when writing)
d %>%
  dplyr::select(ideophone,meaning,language,category,meaning_NL,study,score,rating,C_aspect,C_magnitude,C_modality,C_length,C_irregular,C_cumulative) %>%
  write_xlsx(path="data/ideophones_coded_guessed_rated_simplified.xlsx")

## Figures
Here we generate some descriptive stats and create the main figures for the paper.

```{r descriptive_stats}

d %>%
  group_by(study) %>%
  summarize(max=max(score),min=min(score),mean=mean.na(score))

### Figure: Overview panel
The overview panel is created separately but we use the data from the guessability studies as a canvas.

```{r figures_paper_guessability_canvas}

# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
dp <- d
dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)

# blank plot can be useful as a canvas 
ggplot(data=dp, aes(x=study,y=score)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
  stat_summary(fun=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
  geom_dotplot(colour="white",fill="black",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")

ggsave("figures\\paper-guessability-blank.pdf",height=5,width=7.5)

d %>% filter(filename=="Korean_23_Motion_org") %>% View()

### Figure: Guessability by study and iconicity rating

```{r figures_paper_guessability_rating}
tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
dp <- d
dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)

# panel A: scores by study and cumulative iconicity
pA <- ggplot(data=dp, aes(x=study,y=score,fill=study,colour=study)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
  labs(title="Guessability by study",
       x="study",
       y="guessability (% correct)") +
  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center") +
  NULL

pB <- ggplot(data=dp, aes(x=rating,y=score,fill=study,colour=study)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") +
  labs(title="Guessability (both studies) by iconicity rating",
       x="iconicity rating",
       y="") +
  geom_point(size=2.0,shape=21) +
  geom_smooth(method="lm") +
  NULL

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures\\paper_panelAB_guessing_rating.png",height=5,width=9)

### Figure: Guessability and cumulative iconicity

```{r figures_paper_guessability_coding}

# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
dp <- d
dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)

# panel A: scores by study and cumulative iconicity
pA <- ggplot(data=dp, aes(x=study,y=score,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
  labs(title="Guessability by study",
       x="study",
       y="guessability (% correct)") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white") +
  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center") +
  NULL

pB <- ggplot(data=dp, aes(x=C_cumulative,y=score_z,colour=C_cumulative)) +
  theme_tufte() + theme(legend.position="none") +
  labs(title="Guessability (both studies) by cumulative iconicity",
       x="cumulative iconicity",
       y=expression('guessability ('~italic(z)~')')) +
  geom_half_boxplot() +
  geom_half_point() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  NULL

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures\\paper_panelAB.png",height=5,width=9)

### Figure: Ratings and cumulative iconicity

```{r ratings_and_C}

# panel A: ratings by study and cumulative iconicity
pA <- ggplot(data=dp, aes(x="",y=rating,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
  labs(title="Iconicity ratings for 239 ideophones",
       x="all ideophones",
       y="mean iconicity rating") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  geom_beeswarm(cex=2.5) +
  NULL

pB <- ggplot(data=dp, aes(x=C_cumulative,y=rating,colour=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") +
  labs(title="... by independently coded structural correspondences",
       x="cumulative iconicity",
       y="mean iconicity rating") +
  geom_half_boxplot() +
  geom_half_point() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  NULL

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures\\paper_panelAB_ratings.png",height=5,width=9)

>>>>

# get ratings data
d.ratings <- read_xlsx("data/ideophones_rated_means.xlsx") %>%
  dplyr::select(-category,-list,-item)
d <- left_join(d,d.ratings,by=c("filename","study","language"))

@@
